Activity monitoring analysis
==================================
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

```{r, message=F, warning=F}
library(doBy)
library(chron)
library(ggplot2)
act = read.csv("activity.csv", header = T)
summary = summaryBy(steps ~ date, data = act, FUN = list(sum, mean))

names(summary) = c("date", "steps_Sum", "steps_Mean")
summary[,1:2]
```

Following is the histogram of the number of steps taken

```{r, echo=FALSE}
hist(summary$steps_Sum, xlab = "Number of Steps", main = "Histogram of number of steps taken per day")
meanSteps = mean(summary$steps_Sum, na.rm = T)

```



Following is the mean of the number of steps taken per day is `r format(meanSteps, digits = 6)`
and the median of the values is `r format(median(summary$steps_Sum, na.rm = T), digits = 6)`

```{r,message=F, warning=F,echo=FALSE}
# act$interval = as.factor(act$interval)

timeSummary = aggregate(steps ~ interval, data = act, FUN = mean)

g = ggplot(data = timeSummary, aes(x = interval))
g + geom_bar(colour = "black", fill = "blue") 

```

Following is the number of NA values in the dataset is `r sum(is.na(act$steps))`

```{r, message=F, warning=F,echo=FALSE}
# act$steps[is.na(act$steps)] = timeSummary$steps[timeSummary$interval==act$interval]
act$steps[is.na(act$steps)] = summary$steps_Mean[summary$date==act$date]
summaryNew = summaryBy(steps ~ date, data = act, FUN = sum)
names(summaryNew) = c("date", "steps_Sum")
hist(summaryNew$steps_Sum, xlab = "Number of Steps", main = "Histogram of number of steps taken per day")
meanSteps = mean(summaryNew$steps_Sum, na.rm = T)
```

To substitute the missing values, we use the average steps taken per day(with the NA values) and put them into the previously NA fields in the steps column.

Following is the mean of the number of steps taken per day after accounting for the NA values `r format(meanSteps, digits = 6)`
and the median of the values is `r format(median(summary$steps_Sum, na.rm = T), digits = 6)`

No we take a look at the difference in steps taken considering the day being a weekend or weekday.

```{r, message=F, warning=F, echo=F}
act$weekday = !chron :: is.weekend(act$date)
act$weekday[act$weekday==TRUE] = "Weekday"
act$weekday[act$weekday==FALSE] = "Weekend"
timeSummary = aggregate(steps ~ interval + weekday, data = act, FUN = mean)
g = ggplot(data = timeSummary, aes(x = interval, fill = weekday))
g + geom_bar()
```

